### 2022-09-14 Resources/LogJournal.Rmd
first commit

---
title: "Log"
author: "J.A. Busker"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
```

<!-- Copyright (c) 2022 John Busker. -->
<!-- Licensed under GPLv3. See gpl.md -->

---

The data has been retrived from [uci.](https://archive.ics.uci.edu/ml/datasets/Yeast)

Attributes information:

1.  Sequence Name: Accession number for the SWISS-PROT database  
2.  mcg: McGeoch's method for signal sequence recognition.  
3.  gvh: von Heijne's method for signal sequence recognition.  
4.  alm: Score of the ALOM membrane spanning region prediction program.  
5.  mit: Score of discriminant analysis of the amino acid content of the N-terminal region (20 residues long) of mitochondrial and non-mitochondrial proteins.  
6.  erl: Presence of "HDEL" substring (thought to act as a signal for retention in the endoplasmic reticulum lumen). Binary attribute.  
7.  pox: Peroxisomal targeting signal in the C-terminus.  
8.  vac: Score of discriminant analysis of the amino acid content of vacuolar and extracellular proteins.  
9.  nuc: Score of discriminant analysis of nuclear localization signals of nuclear and non-nuclear proteins.  
10. Class Distribution: The class is the localization site. Consisting of:

        CYT (cytosolic or cytoskeletal)                    463  
        NUC (nuclear)                                      429  
        MIT (mitochondrial)                                244  
        ME3 (membrane protein, no N-terminal signal)       163  
        ME2 (membrane protein, uncleaved signal)            51  
        ME1 (membrane protein, cleaved signal)              44  
        EXC (extracellular)                                 37  
        VAC (vacuolar)                                      30  
        POX (peroxisomal)                                   20  
        ERL (endoplasmic reticulum lumen)                    5

```{r}
data <- as_tibble(read.table("yeast.data", sep = ""))
```