### 2022-09-27 Paper.Rmd
The first commit of the paper

---
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyr)
library(knitr)
library(kableExtra)
library(pander)
library(ggplot2)
library(gridExtra)
```
# Result

To create a great dataset, changes need to be made to the downloaded dataset. The dataset has been gathered of 1484 yeast sequences from SWISS-PROT using the annotations from YPD. In this section, we are going to talk about the results we found from the dataset.

## Codebook

As it is a large dataset, it is wise to look at the codebook first. However, there is no codebook. Nevertheless, there is a file with more information on all the attributes. This is where the information is extracted to make your own codebook.

Below a table with a description of all the attributes abbreviation, explanation and data types.

\begin{center}
Table 1: *The codebook*
\end{center}

```{r}
attr_names <- c("seq.name", "mcg", "gvh", "alm", "mit", "erl",
                "pox", "vac","nuc", "loc.site" )
data_types <- c("str", "float", "float", "float", "float", "float",
                "logical", "float", "float", "factor")
data_labels <- c("Accession number for the SWISS-PROT database",
                 "McGeoch's method for signal sequence recognition",
                 "von Heijne's method for signal sequence recognition",
                 "Score of the ALOM membrane spanning region prediction program",
                 "Score of discriminant analysis of the amino 
                 acid content of the N-terminal region",
                 "Presence of 'HDEL' substring",
                 "Peroxisomal targeting signal in the C-terminus",
                 "Score of discriminant analysis of the amino acid content 
                 of vacuolar and extracellular proteins",
                 "Score of discriminant analysis of nuclear 
                 localization signals of nuclear and non-nuclear proteins",
                 "The class is the localization site")
codebook <- data.frame(Name=attr_names, 
                       Fullname=data_labels, 
                       Datatypes=data_types)
kbl(codebook, booktabs = T, align = "c") %>%
  kable_styling(full_width = T) %>%
  column_spec(2, width = "10cm")
```

The last column is the sequence's localization site. There are 10 different alternatives for this. Here are the 10 in question: 
5  
\begin{center}
Table 2: *Sequence localization sites*
\end{center}

```{r}
abbrv <- c("CYT", "NUC", "MIT", "ME3", "ME2", "ME1", "EXC", "VAC", "POX", "ERL")
loc.fullname <- c("cytosolic or cytoskeletal", "nuclear", "mitochondrial", "membrane protein, no N-terminal signal", "membrane protein, uncleaved signal", "membrane protein, cleaved signal", "extracellular", "vacuolar", "peroxisomal", "endoplasmic reticulum lumen")
amount <- c(463, 429, 244, 163, 51, 44, 37, 30, 20, 5)

pander(data.frame(Abbreviation=abbrv, 
                  Fullname=loc.fullname, 
                  Amount=amount))
```

There are alot of CYT and NUC localization. ERL localization is the least. There are only 5 of these in the dataset.

## Transform data

The data has to be transformed before it can be worked with.